\textbf{The statistical analysis of non-linear longitudinal data in biomedical research using generalized additive models}

### 2021-05-13 Manuscripts/Bookdown testing/Full_document.tex
Need to work on Appendix and Conclusion.

  pdfauthor={Ariel Mundo; Timothy J. Muldoon; John R. Tipton},

@@

\usepackage{subfig}
\usepackage{breqn}
\usepackage[font={small}]{caption}
\usepackage{float}

\title{\textbf{The statistical analysis of non-linear longitudinal data in biomedical research using generalized additive models}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{\emph{Beyond repeated measures ANOVA and Linear Mixed Models}}
\author{Ariel Mundo\footnote{Department of Biomedical Engineering, University of Arkansas, Fayetteville} \and Timothy J. Muldoon\footnote{Department of Biomedical Engineering, University of Arkansas, Fayetteville} \and John R. Tipton\footnote{Department of Mathematical Sciences, University of Arkansas, Fayetteville}}
\date{}

@@

\begin{figure}[!h]
\includegraphics{Manuscript_AM_v5_files/figure-latex/linear-cases-1} \caption{**Simulated linear responses from two groups with correlated (top row) or independent (bottom row) errors using a rm-ANOVA model. A, C:Simulated data with known mean response (linear or quadratic, thin lines) and individual responses (points) showing the dispersion of the data. B,D: Estimations from the rm-ANOVA model for the mean group response (linear of quadratic). Thick lines are the predicted mean response per group, thin lines are the random effects for each subject and points represent the original raw data. The rm-ANOVA model does not pick the trend of the quadratic data**}\label{fig:linear-cases}
\end{figure}

@@

\begin{figure}[!h]
\includegraphics{Manuscript_AM_v5_files/figure-latex/quadratic-cases-1} \caption{**Simulated quadratic responses from two groups with a rm-ANOVA model fitted. A,E:Simulated data with known mean response (lines) and individual responses (points) showing the dispersion of the data. B,F: Generated errors showing the difference in the behavior of correlated and independent errors.  C,G: Simulated known response per group (thick lines) with individual trajectories (thin lines), note that subjects with observations in the area above the mean response tend to stay in that region through the timeline. D,H: Estimations from the rm-ANOVA model for the mean group response. Thick lines are the predicted mean response per group, thin lines are the random effects for each subject and points represent the original raw data.**}\label{fig:quadratic-cases}
\end{figure}

@@

@@

>>>>

@@

\usepackage{lineno}
\usepackage{authblk}

\linenumbers

\author[1]{Ariel I. Mundo}
\author[2]{John R. Tipton}
\author[1]{Timothy J. Muldoon}
\renewcommand\Authands {} %added because an additional "and" appeared after last author
\affil[1]{\footnotesize Department of Biomedical Engineering, University of Arkansas, Fayetteville, AR, USA}
\affil[2]{\footnotesize Department of Mathematical Sciences, University of Arkansas, Fayetteville, AR, USA}
\usepackage{placeins}
\usepackage{subfig}
\usepackage{breqn}
\usepackage[font={small}]{caption}
\usepackage{float}

\title{\textbf{The statistical analysis of non-linear longitudinal data in biomedical research using generalized additive models}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{\emph{Beyond repeated measures ANOVA and Linear Mixed Models}}
\author{}
\date{\vspace{-2.5em}}

@@

\begin{figure}[!h]
\includegraphics{Full_document_files/figure-latex/linear-cases-1} \caption{**Simulated linear responses from two groups with correlated (top row) or independent (bottom row) errors using a rm-ANOVA model. A, C:Simulated data with known mean response (linear or quadratic, thin lines) and individual responses (points) showing the dispersion of the data. B,D: Estimations from the rm-ANOVA model for the mean group response (linear of quadratic). Thick lines are the predicted mean response per group, thin lines are the random effects for each subject and points represent the original raw data. The rm-ANOVA model does not pick the trend of the quadratic data**}\label{fig:linear-cases}
\end{figure}

@@

\begin{figure}[!h]
\includegraphics{Full_document_files/figure-latex/quadratic-cases-1} \caption{**Simulated quadratic responses from two groups with a rm-ANOVA model fitted. A,E:Simulated data with known mean response (lines) and individual responses (points) showing the dispersion of the data. B,F: Generated errors showing the difference in the behavior of correlated and independent errors.  C,G: Simulated known response per group (thick lines) with individual trajectories (thin lines), note that subjects with observations in the area above the mean response tend to stay in that region through the timeline. D,H: Estimations from the rm-ANOVA model for the mean group response. Thick lines are the predicted mean response per group, thin lines are the random effects for each subject and points represent the original raw data.**}\label{fig:quadratic-cases}
\end{figure}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{#generate the response}
\NormalTok{n_time =}\StringTok{ }\DecValTok{6}
\NormalTok{ x <-}\StringTok{ }\KeywordTok{seq}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{6}\NormalTok{, }\DataTypeTok{length.out =}\NormalTok{ n_time)}
\NormalTok{ mu <-}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DecValTok{0}\NormalTok{, }\KeywordTok{length}\NormalTok{(x), }\DecValTok{2}\NormalTok{)}
\NormalTok{ mu[, }\DecValTok{1}\NormalTok{] <-}\StringTok{  }\OperatorTok{-}\NormalTok{(}\FloatTok{0.25} \OperatorTok{*}\StringTok{ }\NormalTok{x}\OperatorTok{^}\DecValTok{2}\NormalTok{) }\FloatTok{+1.5}\OperatorTok{*}\NormalTok{x}\FloatTok{-1.25} \CommentTok{#mean response}
\NormalTok{ mu[, }\DecValTok{2}\NormalTok{] <-}\StringTok{ }\NormalTok{(}\FloatTok{0.25} \OperatorTok{*}\StringTok{ }\NormalTok{x}\OperatorTok{^}\DecValTok{2}\NormalTok{) }\FloatTok{-1.5}\OperatorTok{*}\NormalTok{x}\FloatTok{+1.25} \CommentTok{#mean response}
\NormalTok{ y <-}\StringTok{ }\KeywordTok{array}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DataTypeTok{dim =} \KeywordTok{c}\NormalTok{(}\KeywordTok{length}\NormalTok{(x), }\DecValTok{2}\NormalTok{, }\DecValTok{10}\NormalTok{))}
\NormalTok{ errors <-}\StringTok{ }\KeywordTok{array}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DataTypeTok{dim =} \KeywordTok{c}\NormalTok{(}\KeywordTok{length}\NormalTok{(x), }\DecValTok{2}\NormalTok{, }\DecValTok{10}\NormalTok{))}
 \ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\DecValTok{2}\NormalTok{) \{     }\CommentTok{# number of treatments}
     \ControlFlowTok{for}\NormalTok{ (j }\ControlFlowTok{in} \DecValTok{1}\OperatorTok{:}\DecValTok{10}\NormalTok{) \{  }\CommentTok{# number of subjects}
         \CommentTok{# compound symmetry errors}
\NormalTok{         errors[, i, j] <-}\StringTok{ }\KeywordTok{rmvn}\NormalTok{(}\DecValTok{1}\NormalTok{, }\KeywordTok{rep}\NormalTok{(}\DecValTok{0}\NormalTok{, }\KeywordTok{length}\NormalTok{(x)), }\FloatTok{0.1} \OperatorTok{*}\StringTok{ }\KeywordTok{diag}\NormalTok{(}\DecValTok{6}\NormalTok{) }\OperatorTok{+}\StringTok{ }\FloatTok{0.25} \OperatorTok{*}\StringTok{ }\KeywordTok{matrix}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{6}\NormalTok{, }\DecValTok{6}\NormalTok{))}
\NormalTok{         y[, i, j] <-}\StringTok{ }\NormalTok{mu[, i] }\OperatorTok{+}\StringTok{ }\NormalTok{errors[, i, j]}
\NormalTok{     \}}
\NormalTok{ \}}
  \KeywordTok{dimnames}\NormalTok{(y) <-}\StringTok{ }\KeywordTok{list}\NormalTok{(}\DataTypeTok{time =}\NormalTok{ x, }\DataTypeTok{treatment =} \DecValTok{1}\OperatorTok{:}\DecValTok{2}\NormalTok{, }\DataTypeTok{subject =} \DecValTok{1}\OperatorTok{:}\DecValTok{10}\NormalTok{)}
 \KeywordTok{dimnames}\NormalTok{(errors) <-}\StringTok{ }\KeywordTok{list}\NormalTok{(}\DataTypeTok{time =}\NormalTok{ x, }\DataTypeTok{treatment =} \DecValTok{1}\OperatorTok{:}\DecValTok{2}\NormalTok{, }\DataTypeTok{subject =} \DecValTok{1}\OperatorTok{:}\DecValTok{10}\NormalTok{)}
 \KeywordTok{dimnames}\NormalTok{(mu) <-}\StringTok{ }\KeywordTok{list}\NormalTok{(}\DataTypeTok{time =}\NormalTok{ x, }\DataTypeTok{treatment =} \DecValTok{1}\OperatorTok{:}\DecValTok{2}\NormalTok{)}
\NormalTok{ dat <-}\StringTok{ }\KeywordTok{as.data.frame.table}\NormalTok{(y, }\DataTypeTok{responseName =} \StringTok{"y"}\NormalTok{)}
\NormalTok{ dat_errors <-}\StringTok{ }\KeywordTok{as.data.frame.table}\NormalTok{(errors, }\DataTypeTok{responseName =} \StringTok{"errors"}\NormalTok{)}
\NormalTok{ dat_mu <-}\StringTok{ }\KeywordTok{as.data.frame.table}\NormalTok{(mu, }\DataTypeTok{responseName =} \StringTok{"mu"}\NormalTok{)}
\NormalTok{ dat <-}\StringTok{ }\KeywordTok{left_join}\NormalTok{(dat, dat_errors, }\DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"time"}\NormalTok{, }\StringTok{"treatment"}\NormalTok{, }\StringTok{"subject"}\NormalTok{))}
\NormalTok{ dat <-}\StringTok{ }\KeywordTok{left_join}\NormalTok{(dat, dat_mu, }\DataTypeTok{by =} \KeywordTok{c}\NormalTok{(}\StringTok{"time"}\NormalTok{, }\StringTok{"treatment"}\NormalTok{))}
\NormalTok{ dat}\OperatorTok{$}\NormalTok{time <-}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{(}\KeywordTok{as.character}\NormalTok{(dat}\OperatorTok{$}\NormalTok{time))}

 \CommentTok{#add subject column}
\NormalTok{ dat <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
\StringTok{     }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{subject =} \KeywordTok{factor}\NormalTok{(}\KeywordTok{paste}\NormalTok{(subject, treatment, }\DataTypeTok{sep =} \StringTok{"-"}\NormalTok{)))}

\NormalTok{  dat<-}\KeywordTok{subset}\NormalTok{(dat,treatment}\OperatorTok{==}\DecValTok{1}\NormalTok{)}
\NormalTok{  dat<-dat[,}\KeywordTok{c}\NormalTok{(}\StringTok{'y'}\NormalTok{,}\StringTok{'time'}\NormalTok{)]}
\NormalTok{gm<-}\KeywordTok{gam}\NormalTok{(y}\OperatorTok{~}\KeywordTok{s}\NormalTok{(time,}\DataTypeTok{k=}\DecValTok{5}\NormalTok{),}\DataTypeTok{data=}\NormalTok{dat)}
\NormalTok{model_matrix<-}\KeywordTok{as.data.frame}\NormalTok{(}\KeywordTok{predict}\NormalTok{(gm,}\DataTypeTok{type=}\StringTok{'lpmatrix'}\NormalTok{))}

\NormalTok{x_new <-}\StringTok{ }\NormalTok{(}\KeywordTok{seq}\NormalTok{(}\DecValTok{0}\NormalTok{, }\KeywordTok{max}\NormalTok{(dat}\OperatorTok{$}\NormalTok{time), }\DataTypeTok{length.out =} \DecValTok{100}\NormalTok{))}
\NormalTok{y_pred <-}\StringTok{ }\KeywordTok{predict}\NormalTok{(gm, }\DataTypeTok{newdata=}\NormalTok{dat) }

\NormalTok{time<-}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{6}\NormalTok{)}

\NormalTok{basis<-model_matrix[}\DecValTok{1}\OperatorTok{:}\DecValTok{6}\NormalTok{,] }\CommentTok{#extracting basis}
\CommentTok{#basis<-model_matrix[1:6,-1] #extracting basis}
\KeywordTok{colnames}\NormalTok{(basis)[}\KeywordTok{colnames}\NormalTok{(basis)}\OperatorTok{==}\StringTok{"(Intercept)"}\NormalTok{]<-}\StringTok{"s(time).0"}
\NormalTok{basis<-basis }\OperatorTok{%>%}\StringTok{ }\CommentTok{#pivoting to long format}
\StringTok{  }\KeywordTok{pivot_longer}\NormalTok{(}
    \DataTypeTok{cols=}\KeywordTok{starts_with}\NormalTok{(}\StringTok{"s"}\NormalTok{)}
\NormalTok{  )}\OperatorTok{%>%}
\StringTok{  }\KeywordTok{arrange}\NormalTok{(name) }\CommentTok{#ordering}

\CommentTok{#length of dataframe to be created: number of knots by number of timepoints (minus 1 for the intercept that we won't plot)}
\NormalTok{ln<-}\DecValTok{6}\OperatorTok{*}\NormalTok{(}\KeywordTok{length}\NormalTok{(}\KeywordTok{coef}\NormalTok{(gm))) }

\NormalTok{basis_plot<-}\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{Basis=}\KeywordTok{integer}\NormalTok{(ln),}
                       \DataTypeTok{value_orig=}\KeywordTok{double}\NormalTok{(ln),}
                       \DataTypeTok{time=}\KeywordTok{integer}\NormalTok{(ln),}
                       \DataTypeTok{cof=}\KeywordTok{double}\NormalTok{(ln)}
\NormalTok{)}

\NormalTok{basis_plot}\OperatorTok{$}\NormalTok{time<-}\KeywordTok{rep}\NormalTok{(time) }\CommentTok{#pasting timepoints}
\NormalTok{basis_plot}\OperatorTok{$}\NormalTok{Basis<-}\KeywordTok{factor}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{5}\NormalTok{),}\DataTypeTok{each=}\DecValTok{6}\NormalTok{)) }\CommentTok{#pasting basis number values}
\NormalTok{basis_plot}\OperatorTok{$}\NormalTok{value_orig<-basis}\OperatorTok{$}\NormalTok{value }\CommentTok{#pasting basis values}
\NormalTok{basis_plot}\OperatorTok{$}\NormalTok{cof<-}\KeywordTok{rep}\NormalTok{(}\KeywordTok{coef}\NormalTok{(gm)[}\DecValTok{1}\OperatorTok{:}\DecValTok{5}\NormalTok{],}\DataTypeTok{each=}\DecValTok{6}\NormalTok{) }\CommentTok{#pasting coefficients}
\NormalTok{basis_plot<-basis_plot}\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{mod_val=}\NormalTok{value_orig}\OperatorTok{*}\NormalTok{cof)}

\CommentTok{#creating labeller to change the labels in the basis plots}

\NormalTok{basis_names<-}\KeywordTok{c}\NormalTok{(}
  \StringTok{`}\DataTypeTok{1}\StringTok{`}\NormalTok{=}\StringTok{"Intercept"}\NormalTok{,}
  \StringTok{`}\DataTypeTok{2}\StringTok{`}\NormalTok{=}\StringTok{"1"}\NormalTok{,}
  \StringTok{`}\DataTypeTok{3}\StringTok{`}\NormalTok{=}\StringTok{"2"}\NormalTok{,}
  \StringTok{`}\DataTypeTok{4}\StringTok{`}\NormalTok{=}\StringTok{"3"}\NormalTok{,}
  \StringTok{`}\DataTypeTok{5}\StringTok{`}\NormalTok{=}\StringTok{"4"}
\NormalTok{)}

\CommentTok{#calculating the final spline by aggregating the basis functions}

\NormalTok{spl<-basis_plot}\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(time)}\OperatorTok{%>%}
\StringTok{  }\KeywordTok{summarize}\NormalTok{(}\DataTypeTok{spl=}\KeywordTok{sum}\NormalTok{(mod_val))}

\CommentTok{#original basis}
\NormalTok{sz<-}\DecValTok{1}
\NormalTok{p11<-}\KeywordTok{ggplot}\NormalTok{(basis_plot,}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{time,}\DataTypeTok{y=}\NormalTok{value_orig,}\DataTypeTok{colour=}\KeywordTok{as.factor}\NormalTok{(Basis)))}\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_line}\NormalTok{(}\DataTypeTok{size=}\NormalTok{sz,}\DataTypeTok{show.legend=}\OtherTok{FALSE}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{size=}\NormalTok{sz}\OperatorTok{+}\DecValTok{1}\NormalTok{,}\DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{y=}\StringTok{'Basis functions'}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{Basis,}\DataTypeTok{labeller =} \KeywordTok{as_labeller}\NormalTok{(basis_names))}\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_classic}\NormalTok{()}\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_color_aaas}\NormalTok{()}

\CommentTok{#penalized basis}
\NormalTok{p12<-}\KeywordTok{ggplot}\NormalTok{(basis_plot,}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{time,}\DataTypeTok{y=}\NormalTok{mod_val,}\DataTypeTok{colour=}\KeywordTok{as.factor}\NormalTok{(Basis)))}\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_line}\NormalTok{(}\DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{,}\DataTypeTok{size=}\NormalTok{sz)}\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{,}\DataTypeTok{size=}\NormalTok{sz}\OperatorTok{+}\DecValTok{1}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{y=}\StringTok{'Penalized }\CharTok{\textbackslash{}n}\StringTok{ basis functions'}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_y_continuous}\NormalTok{(}\DataTypeTok{breaks=}\KeywordTok{seq}\NormalTok{(}\OperatorTok{-}\DecValTok{1}\NormalTok{,}\DecValTok{1}\NormalTok{,}\DecValTok{1}\NormalTok{))}\OperatorTok{+}
\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{Basis,}\DataTypeTok{labeller=}\KeywordTok{as_labeller}\NormalTok{(basis_names))}\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_classic}\NormalTok{()}\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_color_aaas}\NormalTok{()}

\CommentTok{#heatmap of the penalization coefficient}
\NormalTok{x_labels<-}\KeywordTok{c}\NormalTok{(}\StringTok{"Intercept"}\NormalTok{,}\StringTok{"1"}\NormalTok{,}\StringTok{"2"}\NormalTok{,}\StringTok{"3"}\NormalTok{,}\StringTok{"4"}\NormalTok{)}
\NormalTok{p13<-}\KeywordTok{ggplot}\NormalTok{(basis_plot,}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{Basis,}\DataTypeTok{y=}\NormalTok{Basis,}\DataTypeTok{fill=}\NormalTok{cof))}\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_tile}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{color=}\StringTok{'black'}\NormalTok{),}\DataTypeTok{size=}\NormalTok{sz}\OperatorTok{+}\DecValTok{1}\NormalTok{,}\DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_tile}\NormalTok{(}\DataTypeTok{size=}\NormalTok{sz}\OperatorTok{+}\DecValTok{1}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_fill_gradient}\NormalTok{(}\DataTypeTok{low =} \StringTok{"white"}\NormalTok{, }\DataTypeTok{high =} \StringTok{"red"}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{x=}\StringTok{'Basis'}\NormalTok{,}\DataTypeTok{y=}\StringTok{'Basis'}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_x_discrete}\NormalTok{(}\DataTypeTok{labels=}\NormalTok{x_labels)}\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_text}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label=}\KeywordTok{round}\NormalTok{(cof,}\DecValTok{2}\NormalTok{)),}\DataTypeTok{size=}\DecValTok{10}\NormalTok{,}\DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_classic}\NormalTok{()}\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.title =} \KeywordTok{element_blank}\NormalTok{())}

\CommentTok{#plotting simulated datapoints and smooth term}
\NormalTok{p14<-}\KeywordTok{ggplot}\NormalTok{(}\DataTypeTok{data=}\NormalTok{dat,}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{time,}\DataTypeTok{y=}\NormalTok{y))}\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{size=}\NormalTok{sz}\OperatorTok{+}\DecValTok{1}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_color_aaas}\NormalTok{()}\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{y=}\StringTok{'Simulated }\CharTok{\textbackslash{}n}\StringTok{ response'}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_line}\NormalTok{(}\DataTypeTok{data=}\NormalTok{spl,}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x=}\NormalTok{time,}\DataTypeTok{y=}\NormalTok{spl),}\DataTypeTok{color=}\StringTok{"#B15731"}\NormalTok{,}\DataTypeTok{size=}\NormalTok{sz}\OperatorTok{+}\DecValTok{1}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_classic}\NormalTok{()}

\CommentTok{#Combining all}
\NormalTok{b_plot<-p11}\OperatorTok{+}\NormalTok{p13}\OperatorTok{+}\NormalTok{p12}\OperatorTok{+}\NormalTok{p14}\OperatorTok{+}\KeywordTok{plot_annotation}\NormalTok{(}\DataTypeTok{tag_levels=}\StringTok{'A'}\NormalTok{)}\OperatorTok{&}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}
     \DataTypeTok{text=}\KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size=}\DecValTok{18}\NormalTok{)}
\NormalTok{     )}

\NormalTok{b_plot}
\end{Highlighting}
\end{Shaded}

\includegraphics{Full_document_files/figure-latex/basis-plot-Appendix-1}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{dat<-}\KeywordTok{tibble}\NormalTok{(}\DataTypeTok{StO2=}\KeywordTok{c}\NormalTok{(}\DecValTok{4}\NormalTok{,}\DecValTok{27}\NormalTok{,}\DecValTok{3}\NormalTok{,}\DecValTok{2}\NormalTok{,}\FloatTok{0.5}\NormalTok{,}\DecValTok{7}\NormalTok{,}\DecValTok{4}\NormalTok{,}\DecValTok{50}\NormalTok{,}\DecValTok{45}\NormalTok{,}\DecValTok{56}\NormalTok{),}
            \DataTypeTok{Day=}\KeywordTok{rep}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{5}\NormalTok{,}\DecValTok{7}\NormalTok{,}\DecValTok{10}\NormalTok{),}\DataTypeTok{times=}\DecValTok{2}\NormalTok{),}
            \DataTypeTok{Group=}\KeywordTok{as.factor}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\StringTok{"Control"}\NormalTok{,}\StringTok{"Treatment"}\NormalTok{),}\DataTypeTok{each=}\DecValTok{5}\NormalTok{))}
\NormalTok{            )}

\CommentTok{## plot the mean response}
\NormalTok{f1<-}\KeywordTok{ggplot}\NormalTok{(dat, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ Day, }\DataTypeTok{y =}\NormalTok{ StO2, }\DataTypeTok{color =}\NormalTok{ Group)) }\OperatorTok{+}
\StringTok{    }\KeywordTok{geom_line}\NormalTok{(}\DataTypeTok{size=}\DecValTok{1}\NormalTok{,}\DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{)}\OperatorTok{+}
\StringTok{    }\KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{,}\DataTypeTok{size=}\FloatTok{1.5}\NormalTok{,}\DataTypeTok{alpha=}\FloatTok{0.5}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{y=}\KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(StO[}\DecValTok{2}\NormalTok{],}\StringTok{' (real)'}\NormalTok{)))}\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_classic}\NormalTok{()}\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_color_aaas}\NormalTok{()}\OperatorTok{+}
\StringTok{    }\KeywordTok{scale_x_continuous}\NormalTok{(}\DataTypeTok{breaks=}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{5}\NormalTok{,}\DecValTok{10}\NormalTok{))}\OperatorTok{+}
\StringTok{    }\KeywordTok{scale_y_continuous}\NormalTok{(}\DataTypeTok{breaks=}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{40}\NormalTok{))}\OperatorTok{+}
\StringTok{  }\KeywordTok{plot_layout}\NormalTok{(}\DataTypeTok{tag_level =} \StringTok{'new'}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}
    \DataTypeTok{plot.background =} \KeywordTok{element_rect}\NormalTok{(}\DataTypeTok{fill =} \StringTok{"transparent"}\NormalTok{, }\DataTypeTok{color =} \OtherTok{NA}\NormalTok{),}
    \DataTypeTok{axis.text=}\KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size=}\DecValTok{14}\NormalTok{)}
\NormalTok{  )}

\CommentTok{#This function simulates data for the tumor data using default parameters of 10 observations per time point,and Standard deviation (sd) of 5%.}
\CommentTok{#Because physiologically StO2 cannot go below 0%, data is  generated with a cutoff value of 0.0001 (the "StO2_sim")}

\NormalTok{simulate_data <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(dat, }\DataTypeTok{n =} \DecValTok{10}\NormalTok{, }\DataTypeTok{sd =} \DecValTok{5}\NormalTok{) \{}
\NormalTok{    dat_sim <-}\StringTok{ }\NormalTok{dat }\OperatorTok{%>%}
\StringTok{        }\KeywordTok{slice}\NormalTok{(}\KeywordTok{rep}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\KeywordTok{n}\NormalTok{(), }\DataTypeTok{each =}\NormalTok{ n)) }\OperatorTok{%>%}
\StringTok{        }\KeywordTok{group_by}\NormalTok{(Group, Day) }\OperatorTok{%>%}
\StringTok{        }\KeywordTok{mutate}\NormalTok{(}
               \DataTypeTok{StO2_sim =} \KeywordTok{pmax}\NormalTok{(}\KeywordTok{rnorm}\NormalTok{(n, StO2, sd), }\FloatTok{0.0001}\NormalTok{),}
               \DataTypeTok{subject=}\KeywordTok{rep}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{10}\NormalTok{),}
               \DataTypeTok{subject=}\KeywordTok{factor}\NormalTok{(}\KeywordTok{paste}\NormalTok{(subject, Group, }\DataTypeTok{sep =} \StringTok{"-"}\NormalTok{))}
\NormalTok{               ) }\OperatorTok{%>%}
\StringTok{        }\KeywordTok{ungroup}\NormalTok{()}

    \KeywordTok{return}\NormalTok{(dat_sim)}
\NormalTok{\}}

\CommentTok{#subject = factor(paste(subject, treatment, sep = "-")))}

\NormalTok{n <-}\StringTok{ }\DecValTok{10} \CommentTok{#number of observations}
\NormalTok{sd <-}\StringTok{ }\DecValTok{10} \CommentTok{#approximate sd from paper}
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{1}\NormalTok{) }\CommentTok{#set seed for reproducibility}
\NormalTok{df <-}\StringTok{ }\DecValTok{6}
\NormalTok{dat_sim <-}\StringTok{ }\KeywordTok{simulate_data}\NormalTok{(dat, n, sd)}

\CommentTok{#plotting simulated data}
\NormalTok{f2<-}\KeywordTok{ggplot}\NormalTok{(dat_sim, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ Day, }\DataTypeTok{y =}\NormalTok{ StO2_sim, }\DataTypeTok{color =}\NormalTok{ Group)) }\OperatorTok{+}
\StringTok{    }\KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{show.legend=}\OtherTok{FALSE}\NormalTok{,}\DataTypeTok{size=}\FloatTok{1.5}\NormalTok{,}\DataTypeTok{alpha=}\FloatTok{0.5}\NormalTok{)}\OperatorTok{+}
\StringTok{    }\KeywordTok{stat_summary}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{y =}\NormalTok{ StO2_sim,}
                     \DataTypeTok{group=}\NormalTok{Group), }
                 \DataTypeTok{fun=}\NormalTok{mean, }\DataTypeTok{geom=}\StringTok{"line"}\NormalTok{,}\DataTypeTok{size=}\DecValTok{1}\NormalTok{,}\DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{)}\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{y=}\KeywordTok{expression}\NormalTok{(}\KeywordTok{atop}\NormalTok{(StO[}\DecValTok{2}\NormalTok{],}\StringTok{'(simulated)'}\NormalTok{)))}\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_classic}\NormalTok{()}\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}
    \DataTypeTok{axis.text=}\KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size=}\DecValTok{22}\NormalTok{)}
\NormalTok{  )}\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_color_aaas}\NormalTok{()}\OperatorTok{+}
\StringTok{    }\KeywordTok{scale_x_continuous}\NormalTok{(}\DataTypeTok{breaks=}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{2}\NormalTok{,}\DecValTok{5}\NormalTok{,}\DecValTok{7}\NormalTok{,}\DecValTok{10}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{#GAM for StO2}
